---
title: "DataTypes"
author: "Helena"
date: "15/06/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Semana 1 - Data Types - Vectors and Lists



```{r exemp1}
x <- c(0.5, 0.6)      #numeric
class(x)

x <- c(TRUE, FALSE)   #logical
class(x)

x <- c(T, F)          #logical
class(x)

x <- c("a", "b", "c") #character
class(x)

x <- 9:29             #integer
class(x)

x <- c(1+0i, 2+4i)    #complex
class(x)



x <- vector("numeric", length=10 )


y <- c(1.7, "a")      #character
class(x)

y <- c(TRUE, 2)       #numeric
class(x)

y <- c("a", TRUE)     #character
class(x)



x <- 0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
as.complex(x)



x <- c("a", "b", "c")
as.numeric(x)
as.logical(x)




x <- list(1, "a", TRUE, 1+4i)
x


```

## Semana 1 - Data Types - Matrices



```{r exemp2}
m <- matrix(nrow = 2, ncol = 3)
m


dim(m)

attributes(m)
```

```{r exemp3}
m <- matrix(data = 1:6, nrow = 2, ncol = 3)
m
```


```{r exemp4}
m <- 1:10
m



dim(m) <- c(2, 5)
m

```
```{r exemp5}
x <- 1:3
y <- 10:12
m <- cbind(x,y)
m


m <- rbind(x,y)
m

```
## Semana 1 - Data Types - Factors



```{r exemp6}
x <- factor(c("yes", "yes", "no", "yes", "no"))
# levels order is alphabetical (defaut)
x


table(x)

unclass(x)


```

```{r exemp7}
x <- factor(c("yes", "yes", "no", "yes", "no"), levels = c("yes", "no"))

x

```

## Semana 1 - Data Types - Missing Values



```{r exemp8}
x <- c(1, 2, NA, 10, 3)
is.na(x)

is.nan(x)

x <- c(1, 2, NaN, NA, 3)
is.na(x)

is.nan(x)
```

## Semana 1 - Data Types - Data Frame

```{r exemp9}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F))
x


nrow(x)

ncol(x)

```

## Semana 1 - Data Types - Names Attribute

```{r exemp10}
x <- 1:3
names(x)

names(x) <- c("foo", "bar", "norf")
x

names(x)

```

```{r exemp11}
x <- list(a = 1, b = 2, c = 3)
x


```


```{r exemp12}
m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("a", "b"), c("c", "d"))
m

```
## Semana 1 - Reading Tabular Data

### Read

- `read.table` or `read.csv` for reading tabular data (cvs - tem cabeçalho e é separado por virgulas)
- `readLines` for reading lines of a text file
- `source` for reading in R code files (`inverse` of `dump`)
- `dget` for reading in R code files (`inverse` of `dput`)
- `load`for reading in saved workspaces
- `unserialize` for reading single R objects in binary form

### Write

- `write.table`
- `write.Lines`
- `dump`
- `dput`
- `save`
- `serialize`

## Semana 1 - Reading Large Tables

```{r exemp13}
initial <- read.table("dataset.txt", nrows = 100)
classes <- sapply(initial, class)
tabAll <- read.table("dataset.txt", colClasses = classes)

#lê os 100 primeiros e vê classes e dai lê tudo

```


## Semana 1 - Textual Data Formats 


dupt - single R object
dump - can be used on multiple R objects

```{r exemp14}
y <- data.frame(a = 1, b = "a")
dput(y)


dput(y,file = "y.R" )
new.y <- dget("y.R")
new.y

```


```{r exemp15}
x <- "foo"
y <- data.frame(a = 1, b = "a")
dump(c("x", "y"), file = "data.R")
rm(x, y)
source("data.R")
y


x


```
## Semana 1 - Connections: Interfaces to the Outside World

- `file` opens a connection to a file
- `gzfile` opens a connection to a file COMPRESSED WITH GZIP
- `bzfile` opens a connection to a file COMPRESSED WITH BZIP2
- `url` opens a connection to a webpage

```{r exemp16}
str(file)

```
- `description` is the name of the file
- `open` is a code indicating
-- "r" read only
-- "w" writing (and initializing new file)
-- "a" appending
-- "rb", "wb", "ab" reading, writing or appending in binary mode (Windows )

```{r exemp17}
con <- file("foo.txt", "r")
data <-read.csv(con)
close(con)

## is the same as:

data <- read.csv("foo.txt")

```




```{r exemp18}
con <- gzfile("words.gz")
x <- readLines(con, 10)
x

```

`writeLines`takes a character vector and writes each element on line at time to a text file


```{r exemp19}
## This might take time

con <- url("http://www.jhsph.edu", "r")
x <- readLines(con, 20)
head(x)

```

## Semana 1 - Subsetting - Basics

```{r exemp20}
x <- c("a", "b", "c", "c", "d", "a")
x[1]

x[2]

x[1:4]

x[x > "a"]

u <- x > "a"
u

x[u]

```

## Semana 1 - Subsetting - Lists

```{r exemp21}
# Subseting a List

x <- list(foo = 1:4, bar = 0.6)
x

x[1] # retorna uma lista com o "nome"

x[[1]]

x$bar

x[["bar"]]

x["bar"]

```

```{r exemp22}
x <- list(foo = 1:4, bar = 0.6, baz = "hello")
x


x[c(1, 3)] #multiplos elementos é um chaves só

```

```{r exemp23}
x <- list(foo = 1:4, bar = 0.6, baz = "hello")
x

name <- "foo"

x[[name]] ##computed index for "foo"

x$name ## elemente "name" dosent exist

x$foo  ## elemment "foo" does exist

```

```{r exemp24}
x <- list(a = list(10, 12, 14), b = c(3.14, 2.8))
x


x[[c(1, 3)]]

x[[1]][[3]]

x[[c(2, 1)]]
```

## Semana 1 - Subsetting - Matrices

```{r exemp25}
x <- matrix(1:6,2,3)
x


x[1, 2]

x[2, 1]

x[1,  ]

x[ , 2]
```

```{r exemp26}
x <- matrix(1:6,2,3)
x


x[1, 2] # resulta num vetor ao inves de uma matriz

x[1, 2, drop = FALSE] # BY DEFAULT ITs true and "drop" 1 dimension (permanece matriz)


x[1,  ] #vetor

x[1,  , drop = FALSE] #matriz


```

## Semana 1 - Subsetting - Partial Matching

```{r exemp27}
x <- list(aardvark = 1:5)
x

x$a

x[["a"]]

x[["a", exact = FALSE]]

```

## Semana 1 - Subsetting - Partial Matching

```{r exemp28}
x <- c(1, 2, NA, 4, NA, 5)
x

bad <- is.na(x)
bad

x[!bad]

good <- !(is.na(x))
good

x[good]

```


So what if there are a multiple vectors or multiple objects and you want to take and each one has a kind of missing values in slightly different places and you kind of, you want to take the subset of all the objects that have no missing values.


```{r exemp29}
x <- c(1, 2, NA, 4, NA, 5)
x
y <- c("a", "b", NA, "d", NA, "f")
y

good <- complete.cases(x, y)
good

x[good]

y[good]

```


```{r exemp30}
x <- c(1, 2, 3, NA, 4, NA, 5)
x
y <- c("a", "b", NA, "d", NA, "f", "g")
y

good <- complete.cases(x, y)
good

x[good]

y[good]

```


```{r exemp31}
airquality[1:6, ]

good <- complete.cases(airquality) #diz quais linhas estão completas. é um vetor
good[1:6]


airquality[good, ][1:6, ]


```


## Semana 1 - Vectorized Operations

### Vetor

```{r exemp32}
x <- 1:4; y <- 6:9
x + y # 1o + 1o, 2o + 2o etc

x > 2

x >= 2

y == 8

x * y

x / y

```

### Matriz

```{r exemp33}
x <- matrix(1:4, 2, 2)
x

y <- matrix(rep(10, 4), 2, 2)
y

x * y #element-wise multiplication

x / y

x %*% y #true matrix multiplication

x %/% y

```

Quiz

```{r exemp34}
x <- 4L
x
class(x)

x <- c(4, TRUE)
x
class(x)


x <- c(1,3, 5)
y <- c(3, 2, 10)
rbind(x, y)

x <- list(2, "a", "b", TRUE)
x
x[[2]]

class(x[[2]])


 x <- 1:4
 y <- 2:3

w <- x + y
w
class(w)
```


```{r exemp35}
library(readr)
hw1_data <- read.csv("hw1_data.csv")

x <- is.na(hw1_data[ , 1])
x

summary(x)

summary(hw1_data)



```

## Semana 2 - Your First R Function

```{r exemp36}
add2 <- function(x, y){
  x + y # retorna automaticamente o resultado da ultima expressão na função
}

# Chamando a função:

add2(3, 5)

```



```{r exemp37}
above10 <- function(x){
  use <- (x > 10)
  x[use]
}

above <- function(x, n){
  use <- (x > n)
  x[use]
}

# Chamando a função:
x <- 1:20
x

above10(x)

above(x, 15)

```


```{r exemp38}

above <- function(x, n = 10){ #se não especificado n = 10
  use <- (x > n)
  x[use]
}

# Chamando a função:
x <- 1:20
x

above(x)


```



```{r exempX}
columean <- function(y) {
  nc <- ncol(y)
  means <- numeric(nc) 
  ## vetor vazio para as medias com num_elementos = num_colunas matriz original
  for (i in 1:nc) {
    means[i] <- mean(y[ , i])
  }
  means
}

# Chamando a função:

columean(airquality)


```


```{r exempX}
columean <- function(y, remNA = TRUE) {
  nc <- ncol(y)
  means <- numeric(nc) 
  ## vetor vazio para as medias com num_elementos = num_colunas matriz original
  for (i in 1:nc) {
    means[i] <- mean(y[ , i], na.rm = remNA )
  }
  means
}

# Chamando a função:

columean(airquality)

columean(airquality, FALSE)

```



```{r exempX}

columean <- function(y, col) {
  nl <- nrow(y)
  count <- 0
  soma <- 0
  for (i in 1:nl) {
    if (y[i , 5] == "virginica"){
      count <- count + 1
      soma <- y[i, col]
  }}
  means <- soma / count
}

z <- columean(iris, 1)
z


w <- apply(iris[,1:4], 2, mean)
w


x <- colMeans(iris)
x

```



```{r exempX}

#pollutantmean <- function(directory = "specdata", pollutant = "sulfate", id = 1:332){

directory = "specdata"
pollutant = "sulfate"
id.inicio = 70
id.fim = 72

## List filenames to be merged. 
filenames <- list.files(path=directory,pattern="*.csv")

## Print filenames to be merged
#print(filenames)

## Full path to csv filenames
fullpath <- file.path(directory,filenames)

## Print Full Path to the files
#print(fullpath)

## Merge listed files from the path above

dataset <- do.call("rbind",lapply(fullpath,FUN=function(files){ read.csv(files)}))

#dataset <- lapply(fullpath, read.csv)

#class(dataset)

print("----DADOS COMPLETOS: ----")

summary(dataset)

print("---- HEAD FILTRO ----")

filtro <- with(dataset, (ID >= id[id.inicio] & ID <= id[id.fim]) )

#head(filtro)

print("----DADOS ID FILTRADO----")

data.filtrada <- dataset$nitrate[filtro]
summary(data.filtrada)

#print("--------")
#by(dataset, dataset$sulfate, summary)

#print("---- JUST GOOD ----")
#good <- complete.cases(dataset)
#good.dataset <- dataset$sulfate[good]
#summary(good.dataset)

print("----MEDIA SEM NA SULFATO (FILTRADO) ----")

mean(data.filtrada, na.rm = TRUE)

#if(pollutant == "sulfate"){
#mean(new_dataset$sulfate, na.rm = TRUE)
#}

## Print the merged csv dataset, if its large use `head()` function to get glimpse of merged dataset
#dataset


#}


# pollutantmean()


```


```{r exempX}

directory = "specdata"
id.inicio = 70
id.fim = 72

## List filenames to be merged. 
filenames <- list.files(path=directory,pattern="*.csv")

## Print filenames to be merged
#print(filenames)

## Full path to csv filenames
fullpath <- file.path(directory,filenames)

## Print Full Path to the files
#print(fullpath)

## Merge listed files from the path above

dataset <- do.call("rbind",lapply(fullpath,FUN=function(files){ read.csv(files)}))

print("----DADOS COMPLETOS: ----")

summary(dataset)

print("Sulfato:")

mean(dataset$sulfate, na.rm = TRUE)

print("Nitrato:")

mean(dataset$nitrate, na.rm = TRUE)

print("---- HEAD FILTRO ----")

filtro <- with(dataset, (ID >= id[id.inicio] & ID <= id[id.fim]) )

head(filtro)

print("---- JUST GOOD ----")
good <- complete.cases(filtro)
good.dataset <- data.filtrada[good]
summary(good.dataset)

#print("----DADOS ID FILTRADO----")

#data.filtrada <- dataset$nitrate[filtro]
#summary(data.filtrada)

print("----MEDIA SEM NA SULFATO (FILTRADO) ----")

mean(good.dataset)

```


```{r exempX}
data.new <- rbind(read.csv("specdata/070.csv"), read.csv("specdata/071.csv"), read.csv("specdata/072.csv"))

summary(data.new)

mean(data.new$nitrate, na.rm = TRUE)



```



```{r exempX}

data.new <- rbind(read.csv("specdata/034.csv"))

summary(data.new)

mean(data.new$sulfate, na.rm = TRUE)


```




```{r exempX}

#cc <- complete("specdata", c(6, 10, 20, 34, 100, 200, 310))
#print(cc$nobs)

data.new2 <- rbind(read.csv("specdata/054.csv"))

filtro2 <- complete.cases(data.new2)

head(filtro2)

dtf2 <- data.new2$sulfate[filtro2]

head(dtf2)

length(dtf2)

print("---- 3 -----")

print(R.version.string)

RNGversion("3.5.1")
#RNGkind(sample.kind = "Rejection")

set.seed(42)
use <- sample(332, 10)
use
data.new3 <- rbind(read.csv("specdata/044.csv"))
good3 <- complete.cases(data.new3)
good.dataset3 <- data.new3$sulfate[good3]
length(good.dataset3)

data.new3 <- rbind(read.csv("specdata/304.csv"))
good3 <- complete.cases(data.new3)
good.dataset3 <- data.new3$sulfate[good3]
length(good.dataset3)

data.new3 <- rbind(read.csv("specdata/311.csv"))
good3 <- complete.cases(data.new3)
good.dataset3 <- data.new3$sulfate[good3]
length(good.dataset3)




```



```{r exempX}

directory = "specdata"

filenames <- list.files(path=directory,pattern="*.csv")

fullpath <- file.path(directory,filenames)

dataset <- do.call("rbind",lapply(fullpath,FUN=function(files){ read.csv(files)}))

limpo <- na.exclude(dataset)

summary(limpo)

#mean(limpo$sulfate, na.rm = FALSE)


correl <- matrix(c(limpo$ID, limpo$nitrate/limpo$sulfate), ncol = 2)

cor

correl


```





```{r w3}


makeVector <- function(x = numeric()) {
        m <- NULL
        set <- function(y) {
                x <<- y
                m <<- NULL
        }
        get <- function() x
        setmean <- function(mean) m <<- mean
        getmean <- function() m
        list(set = set, get = get,
             setmean = setmean,
             getmean = getmean)
}


cachemean <- function(x, ...) {
        m <- x$getmean()
        if(!is.null(m)) {
                message("getting cached data")
                return(m)
        }
        data <- x$get()
        m <- mean(data, ...)
        x$setmean(m)
        m
}






makeCacheMatrix <- function(x = matrix()) {
        m <- NULL
        set <- function(y) {
                x <<- y
                m <<- NULL
        }
        get <- function() x
        setsolve <- function(solve) m <<- solve
        getsolve <- function() m
        list(set = set, get = get,
             setsolve = setsolve,
             getsolve = getsolve)  

}


cacheSolve <- function(x, ...) {
        ## Return a matrix that is the inverse of 'x'
  m <- x$getsolve()
  if(!is.null(m)) {
      message("getting cached data")
      return(m)
  }
  data <- x$get()
  m <- solve(data, ...)
  x$setsolve(m)
  m
  
}



```




``` {r w4}
dataoutcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
head(dataoutcome)
ncol(dataoutcome)


dataoutcome[, 11] <- as.numeric(dataoutcome[, 11])
## You may get a warning about NAs being introduced; that is okay
hist(dataoutcome[, 11])


#best <- function(state, woutcome) {
state <- "AK"
woutcome <- "heart attack"
  
  validstates <- c("AL","AK","AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY")
  validoutcomes <- c("heart attack", "heart failure", "pneumonia")
  
  
  #checkValid(state, validstates)
  #checkValid(woutcome, validoutcomes)
  
  filtro <- dataoutcome[,7] == state
  
  head(filtro)
  
  filtered <- dataoutcome[filtro]
  
  head(filtered)
  
    
## Read outcome data
## Check that state and outcome are valid
## Return hospital name in that state with lowest 30-day death
## rate
#}



```